---
title: "p8105_hw1_lr2854"
author: "Lauren Richter"
date: 2021-09-29
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

This problem is intended to emphasize variable types and introduce coercion; some awareness of how R treats numeric, character, and factor variables is necessary for working with these data types in practice.

Create a data frame comprised of:

- a random sample of size 10 from a standard Normal distribution
- a logical vector indicating whether elements of the sample are greater than 0
- a character vector of length 10
- a factor vector of length 10, with 3 different factor “levels”

```{r prob1a, warning = FALSE}

prob1a_df = tibble(
  norm_sample_10 = rnorm(10),
  is_gt_0 = c(norm_sample_10 > 0),
  char_vec = letters[1:10],
  factor_vec = factor(
    c(rep("cow",4),
      rep("chicken",3),
      rep("dog",3)
      )
    )
)

prob1a_df
```
Try to take the mean of each variable in your dataframe. What works and what doesn’t?

*Hint: to take the mean of a variable in a dataframe, you need to pull the variable out of the dataframe. Try loading the tidyverse and using the ```pull``` function.*

**ANSWER:** You can take the mean of the numeric data, but not the character data (which includes the factor data). The logical vector is converted to numeric 0 or 1 and the mean is taken from that.
```{r prob1b, warning = FALSE}

mean(pull(prob1a_df,norm_sample_10))
mean(pull(prob1a_df,is_gt_0))
mean(pull(prob1a_df,char_vec))
mean(pull(prob1a_df,factor_vec))

```

In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?

**ANSWER:** Boolean statements are converted to numeric ones (which already happened previously...?). The character vector cannot be assigned a number because there is no obvious order. The factor vector is converted from categorical to numeric data, so "cow" becomes 2, "chicken" becomes 1, and "dog" becomes 3 (because it ordered them alphabetically, not by the order in which they appear in the tibble).

```{r prob1c, warning = FALSE, results = 'hide'}

as.numeric(pull(prob1a_df,is_gt_0))
as.numeric(pull(prob1a_df,char_vec))
as.numeric(pull(prob1a_df,factor_vec))

```

## Problem 2

This problem focuses the use of inline R code, plotting, and the behavior of ```ggplot``` for variables of different types.

Use the code below to download the a package containing the ```penguins``` dataset:

```{r penguins, warning = FALSE}
data("penguins", package = "palmerpenguins")
```

Write a short description of the ```penguins``` dataset (not the ```penguins_raw``` dataset) using inline R code, including:

- the data in this dataset, including names / values of important variables
- the size of the dataset (using ```nrow``` and ```ncol```)
- the mean flipper length
- Make a scatterplot of ```flipper_length_mm``` (y) vs ```bill_length_mm``` (x); color points using the ```species``` variable (adding ```color = ...``` inside of ```aes``` in your ggplot code should help).

Export your first scatterplot to your project directory using ```ggsave```.

**ANSWER:** The Penguins dataset is a 344 x 8 dataframe. The variables include the species (Adelie, Chinstrap, and Gentoo), the island (Biscoe, Dream, and Torgersen), the bill length, bill depth, and flipper length in mm, body mass in g, and sex (male or female). You can get the names of the different species or islands using the ```levels``` function.

The mean penguin flipper length is 200.9152, with 2 NA values dropped.

```{r mean_flipper_len}
mean(penguins$flipper_length_mm,na.rm = TRUE)
```

```{r penguins_levels, warning = FALSE, echo = FALSE}
levels(penguins$species)
levels(penguins$island)
levels(penguins$sex)
```

